### 2021-03-04 paper2/3-method.tex
Fix headings and remove duplicate bib entries

\subsection{Conditional Flow}\label{cflow}
\subsubsection{Invertible Neural Networks}
Invertible neural networks are a class of neural network architecture characterised by a bijective mapping between their inputs and output \citep{Dinh2014}. The transformations are designed such that their inverses and Jacobians are efficiently computable.
These flow-based models permit \emph{exact} likelihood estimation \citep{normflows2015} through the warping of a base density with a series of invertible transformations and computing the resulting, highly multi-modal, but still normalised, density, using the change of variable theorem:
% Flow-GAN \cite{grover2018flowgan} combines the \emph{exact} log-likelihood estimation of the invertible network with the adversarial training of a GAN.
\begin{align}
\begin{split}
  \log p(x) &= \log p(z) + 
   \sum \log \left| \det\left( \frac{\diff h_i}{ h_{i-1}}\right) \right|, %\\
  \quad p(z) = \mathcal{N}(z; 0, \mathbb{I})
  \label{eq:changeofvariables}
\end{split}
\end{align}
where $h_i$ refers to the outputs of the layers of the network and $p(z)$ is the base density, specifically an Isotropic Gaussian in our case.
Training of the invertible neural network is then reduced to maximising $\log p(x)$ over the training set,
i.e.\ maximising the probability the network assigns to samples in the training set.

@@

% \textbf{Coupling layers}. Dinh et al. \cite{Dinh2014} introduced a simple yet powerful invertible transformation in the form of coupling layers. Ease of invertibility is achieved by updating only half of the input vector with a function that itself is trivially invertible but that is at the same time parameterised by a arbitrarily complex operation not subject to the invertibility constraint (e.g. a multi-layer neural network).
% Concretely, the vector $\bm{u}$ is split into two evenly sized vectors: $\bm{u} = [\bm{u}_1, \bm{u}_2]$.
% The output of the coupling layer is then a concatenation of vectors $\bm{v}_1$ and $\bm{v}_2$,
% where $\bm{v}_1 = s \cdot \bm{u}_1 + t$ and $\bm{v}_2 = \bm{u}_2$, with the affine parameters $s$ and $t$ generated by a non-invertible function of $\bm{u}_2$.

% \textbf{1x1 Convolutions}. Kingma and Dhariwal \cite{KinDha18} introduced invertible 1x1 convolution as a generalised permutation operation and show that determinant can be computed efficiently using an LU decomposition of the weights.

% \textbf{Spatial downsampling.} To downsample the spatial dimensions and promote mixing between the variables, \cite{Dinh2014} first mask the input with a checkerboard pattern before reshaping (transforming a $c \times h\times w$ tensor into a $4c \times \frac{1}{2} h\times \frac{1}{2}w$). Each "level" of the network is demarcated by a downsampling operation.

\subsubsection{The Benefits of Bijectivity}
Using an invertible network to generate our encoding, $\bm{z}_u$, carries a number of advantages over other approaches.
Ordinarily, the main benefit of flow-based models is that they permit exact density estimation. 
However, since we are not interested in sampling from the model's distribution, in our case the likelihood term serves as a regulariser, as it does for  \citet{JacSmeOya18}. 
Critically, this forces the mean of each latent dimension to zero enabling null-sampling. 
The invertible property of the network guarantees the preservation of all information relevant to $y$ which is independent of $s$, regardless of how it is allocated in the output space.
Secondly, we conjecture that the encodings are more robust to out-of-distribution data.
Whereas an autoencoder could map a previously seen input and a previously unseen input to the same representation,
an invertible network sidesteps this due to the network's bijective property, ensuring all relevant information is stored somewhere. This opens up the possibility of transfer learning between datasets with a similar manifestation of $s$, as we demonstrate in the Appendix~\ref{sec:transfer-learning}.

Under our framework, the invertible network $f$ maps the inputs $\bm{x}$ to a representation $\bm{z}_u$:
$f(\bm{x}) = \bm{z}$.
We interpret the embedding $\bm{z}$ as being the concatenation of two smaller embeddings: $\bm{z} = [\bm{z}_u, \bm{z}_b]$.
The dimensionality of $\bm{z}_b$, and $\bm{z}_u$, by complement, is a free parameter (see Appendix~\ref{ssec:partition-size-matters} for tuning strategies).
As $f$ is invertible, $\bm{x}$ can be recovered like so:
\begin{align}
  \bm{x} = f^{-1}([\bm{z}_u, \bm{z}_b])
  \label{eq:zreconstruct}
\end{align}
where $\bm{z}_b$ is required for equality of the output dimension and input dimension to satisfy the bijectivity of the network -- we cannot output $\bm{z}_u$ alone, but have to output $\bm{z}_b$ as well. In order to generate the pre-image of $\bm{z}_u$, we perform null-sampling with respect to $\bm{z}_b$ by zeroing-out the elements of $\bm{z}_b$ (such that $\bm{x}_{u} = f^{-1}([\bm{z}_{u}, \stackrel{\rightarrow}{0}])$), i.e. setting them to the mean of the prior density, $\mathcal{N}(z;0, I)$.

How can we be sure that $\bm{z}_u$ contains enough information about $y$?
The importance of the invertible architecture bears out from this consideration. %, as the bijectivity of the network guarantees preservation of all information about the input. %
% The existence of the inverse, $f^{-1}$,  p $x$ from $z$ because $f^{-1}$ exists and can do just that.
As long as $\bm{z}_b$ does not contain the information about $y$, $\bm{z}_u$ necessarily must.
We can raise or lower the information capacity of $\bm{z}_b$ by adjusting its size;
this should be set to the smallest size sufficient to capture all information about $s$, so as not to sacrifice class-relevant information.
Appendix~\ref{ssec:size-of-zb} explores the effects of the size further.

@@

% \subsubsection{Tuning the Partition Sizes}

% \subsubsection{The Pitfalls of Adversarial Training}

% \textbf{Preprocessing}.
% Heuristically, we found that preprocessing the data with an autoencoder stabilises and accelerates training of the cFlow model. The autoencoder was pretrained on the pretraining set solely to minimise reconstruction loss and its weights frozen at the time of the INN's training. While this means the INN is not truly lossless with respect to the uncompressed data, its bijectivity is leveraged to ensure semantically-relevant information is not discarded during the pre-training phase, which is still applicable since the autoencoder is not trained jointly with the INN to maximise the adversarial loss. Since the autoencoder is optimised for compression, information about both the spurious and non-spurious attributes is captured impartially in its encoding.

>>>>

\subsection{Conditional Flow}\label{cflow}
\paragraph{Invertible Neural Networks.}
Invertible neural networks are a class of neural network architecture characterised by a bijective mapping between their inputs and output \citep{Dinh2014}. The transformations are designed such that their inverses and Jacobians are efficiently computable.
These flow-based models permit \emph{exact} likelihood estimation \citep{normflows2015} through the warping of a base density with a series of invertible transformations and computing the resulting, highly multi-modal, but still normalised, density, using the change of variable theorem:
% Flow-GAN \cite{grover2018flowgan} combines the \emph{exact} log-likelihood estimation of the invertible network with the adversarial training of a GAN.
\begin{align}
\begin{split}
  \log p(x) &= \log p(z) + 
   \sum \log \left| \det\left( \frac{\diff h_i}{ h_{i-1}}\right) \right|, %\\
  \quad p(z) = \mathcal{N}(z; 0, \mathbb{I})
  \label{eq:changeofvariables}
\end{split}
\end{align}
where $h_i$ refers to the outputs of the layers of the network and $p(z)$ is the base density, specifically an Isotropic Gaussian in our case.
Training of the invertible neural network is then reduced to maximising $\log p(x)$ over the training set,
i.e.\ maximising the probability the network assigns to samples in the training set.

@@

% \paragraph{Coupling layers}. Dinh et al. \cite{Dinh2014} introduced a simple yet powerful invertible transformation in the form of coupling layers. Ease of invertibility is achieved by updating only half of the input vector with a function that itself is trivially invertible but that is at the same time parameterised by a arbitrarily complex operation not subject to the invertibility constraint (e.g. a multi-layer neural network).
% Concretely, the vector $\bm{u}$ is split into two evenly sized vectors: $\bm{u} = [\bm{u}_1, \bm{u}_2]$.
% The output of the coupling layer is then a concatenation of vectors $\bm{v}_1$ and $\bm{v}_2$,
% where $\bm{v}_1 = s \cdot \bm{u}_1 + t$ and $\bm{v}_2 = \bm{u}_2$, with the affine parameters $s$ and $t$ generated by a non-invertible function of $\bm{u}_2$.

% \paragraph{1x1 Convolutions}. Kingma and Dhariwal \cite{KinDha18} introduced invertible 1x1 convolution as a generalised permutation operation and show that determinant can be computed efficiently using an LU decomposition of the weights.

% \paragraph{Spatial downsampling.} To downsample the spatial dimensions and promote mixing between the variables, \cite{Dinh2014} first mask the input with a checkerboard pattern before reshaping (transforming a $c \times h\times w$ tensor into a $4c \times \frac{1}{2} h\times \frac{1}{2}w$). Each "level" of the network is demarcated by a downsampling operation.

\paragraph{The Benefits of Bijectivity.}
Using an invertible network to generate our encoding, $\bm{z}_u$, carries a number of advantages over other approaches.
Ordinarily, the main benefit of flow-based models is that they permit exact density estimation. 
However, since we are not interested in sampling from the model's distribution, in our case the likelihood term serves as a regulariser, as it does for  \citet{JacSmeOya18}. 
Critically, this forces the mean of each latent dimension to zero enabling null-sampling. 
The invertible property of the network guarantees the preservation of all information relevant to $y$ which is independent of $s$, regardless of how it is allocated in the output space.
Secondly, we conjecture that the encodings are more robust to out-of-distribution data.
Whereas an autoencoder could map a previously seen input and a previously unseen input to the same representation,
an invertible network sidesteps this due to the network's bijective property, ensuring all relevant information is stored somewhere. This opens up the possibility of transfer learning between datasets with a similar manifestation of $s$, as we demonstrate in the Appendix~\ref{sec:transfer-learning}.

Under our framework, the invertible network $f$ maps the inputs $\bm{x}$ to a representation $\bm{z}_u$:
$f(\bm{x}) = \bm{z}$.
We interpret the embedding $\bm{z}$ as being the concatenation of two smaller embeddings: $\bm{z} = [\bm{z}_u, \bm{z}_b]$.
The dimensionality of $\bm{z}_b$, and $\bm{z}_u$, by complement, is a free parameter (see section~\ref{sec:optimisation-details} for tuning strategies).
As $f$ is invertible, $\bm{x}$ can be recovered like so:
\begin{align}
  \bm{x} = f^{-1}([\bm{z}_u, \bm{z}_b])
  \label{eq:zreconstruct}
\end{align}
where $\bm{z}_b$ is required for equality of the output dimension and input dimension to satisfy the bijectivity of the network -- we cannot output $\bm{z}_u$ alone, but have to output $\bm{z}_b$ as well. In order to generate the pre-image of $\bm{z}_u$, we perform null-sampling with respect to $\bm{z}_b$ by zeroing-out the elements of $\bm{z}_b$ (such that $\bm{x}_{u} = f^{-1}([\bm{z}_{u}, \stackrel{\rightarrow}{0}])$), i.e. setting them to the mean of the prior density, $\mathcal{N}(z;0, I)$.

How can we be sure that $\bm{z}_u$ contains enough information about $y$?
The importance of the invertible architecture bears out from this consideration. %, as the bijectivity of the network guarantees preservation of all information about the input. %
% The existence of the inverse, $f^{-1}$,  p $x$ from $z$ because $f^{-1}$ exists and can do just that.
As long as $\bm{z}_b$ does not contain the information about $y$, $\bm{z}_u$ necessarily must.
We can raise or lower the information capacity of $\bm{z}_b$ by adjusting its size;
this should be set to the smallest size sufficient to capture all information about $s$, so as not to sacrifice class-relevant information.
Section~\ref{sec:additional-results} explores the effects of the size further.

@@

% \paragraph{Tuning the Partition Sizes.}

% \paragraph{The Pitfalls of Adversarial Training.}

% \paragraph{Preprocessing}.
% Heuristically, we found that preprocessing the data with an autoencoder stabilises and accelerates training of the cFlow model. The autoencoder was pretrained on the pretraining set solely to minimise reconstruction loss and its weights frozen at the time of the INN's training. While this means the INN is not truly lossless with respect to the uncompressed data, its bijectivity is leveraged to ensure semantically-relevant information is not discarded during the pre-training phase, which is still applicable since the autoencoder is not trained jointly with the INN to maximise the adversarial loss. Since the autoencoder is optimised for compression, information about both the spurious and non-spurious attributes is captured impartially in its encoding.

### 2021-03-04 paper2/4-experiments.tex
Fix headings and remove duplicate bib entries

\subsubsection{cMNIST}\label{ssec:cmnist}
The coloured MNIST (cMNIST) dataset is a variant of the MNIST dataset in which the digits are coloured.
In the training set, the colours have a one-to-one correspondence with the digit class.
In the test set (and the representative set), colours are assigned randomly.
The colours are drawn from Gaussians with 10 different means.
We follow the colourisation procedure outlined by~\citet{kim2019learning}, with the mean colour values selected so as to be maximally dispersed.
The full list of such values can be found in Appendix~\ref{sec:color-details}.
We produce multiple variants of the cMNIST dataset corresponding to different standard deviations $\sigma$ for the colour sampling:
$\sigma \in \{0.00, 0.01, ..., 0.05 \}$.
% The images are symmetrically zero-padded to a size of $32 \times 32$ to allow for an additional downsampling stage in the cFlow model.
% No additional data augmentation is used.
% 24,000 of the 36,000 training samples were set aside for representative, the remaining 10,000 for training the downstream classifier.
% We continue to use 10,000 samples usually reserved for testing on vanilla MNIST for that same purpose.

@@

\subsubsection{CelebA}
To evaluate the effectiveness of our framework on real-world image data we use the CelebA dataset~\citep{liu2015faceattributes}, consisting of 202,599 celebrity images.
These images are annotated with various binary physical  attributes, including ``gender'', ``hair color'', ``young'', etc, from which we  select our sensitive and target attributes.
The images are centre cropped and resized to $64\times64$, as is standard practice.
For our experiments, we designate ``gender'' as the sensitive attribute,
and ``smiling'' and ``high cheekbones'' as target attributes.
We chose gender as the sensitive attribute as it a common sensitive attribute in the fairness literature.
For the target attributes, we chose attributes that are harder to learn than gender and which do not correlate too strongly with gender in the dataset
(``wearing lipstick'' for example being an attribute too closely correlated with gender).
% The sampling and evaluation procedure is identical to that conducted for the UCI Adult dataset.
% We evaluate the performance of all models for mixing factors ($\eta$) of value $\{0, 0.1, ..., 1\}$. 
The model is trained on the representative set (normal subset of CelebA)
and is then used to encode the artificially biased training set and the test set.
The results for the most strongly biased training set ($\eta=0$) can be found in Fig.~\ref{fig:celeba-targets}.
Our method outperforms the baselines in accuracy and fairness.

We also assess performance for different mixing factors ($\eta$) which correspond to varying degrees of bias in the training set
(see Fig.~\ref{fig:celeba-multiplot}).
This is to verify that the model does not \emph{harm} performance when there is not much bias in the training set.
For these experiments, the model is trained once on the representative set and is then used to encode different training sets.
The results show that for the intermediate values of $\eta$, our model incurs a small penalty in terms of accuracy,
but at the same time makes the results \emph{fairer} (corresponding to an accuracy-fairness trade-off). Qualitative results can be found in Fig.~\ref{fig:celeba_cflow} (images from cVAE can be found in Appendix~\ref{sec:qual-results-celeba}).

To show that our method can handle multinomial, as well as binary, sensitive attributes, we also conduct experiments with $s=\textrm{hair color}$ as a ternary attribute (``Blonde'', ``Black'', ``Brown''), excluding ``Red'' because of the paucity of samples and the noisiness of their labels. The results for these experiments can be found in Appendix~\ref{ssec:multi-s}.
% Looking at the produced invariant representations,
% The failure of the cVAE to model complex interactions between $x_u$ and $s$ is brought to bear here.
% Due to the balancing between the adversarial and reconstruction losses,
% information relevant to the downstream task of predicting ``smiling'' is lost during the pre-training phase,
% leading to poor performance at all values of $\eta$ except the most extreme ones,
% for which it does reasonably well at mitigating the sampling bias (Fig.~\ref{fig:celeba_chart}).
% Meanwhile, cFlow does not suffer this drawback -- the images remain sharp
% (see Fig.~\ref{fig:cflow_celeba_recon_y} in comparison to Fig. \ref{fig:cvae_celeba_recon_y})
% and the accuracy obtained by training on them is comparable to the baseline at intermediary values -- and at the same time matches and outperforms the cVAE at $\eta = 0$ and $\eta = 1$, respectively.
% For these same values,
% incorporating the mutual information loss proposed by~\cite{ln2l} into the classifier's loss function greatly harmed performance measured by accuracy;
% the DP is zero due to all predictions being negative.

@@

\subsubsection{Results for the UCI Adult dataset}
The UCI Adult dataset consists of census data and is commonly used to evaluate models focused on algorithmic fairness.
Following convention, we designate ``gender'' as the sensitive attribute $s$ and whether an individual's salary is \$50,000 or greater as $y$.
We show the performance of our approach in comparison to baseline approaches in Fig. \ref{fig:adult-chart}.
We evaluate the performance of all models for mixing factors ($\eta$) $0$ and $1$. 
Results shown in Fig. \ref{fig:adult-chart} show that
%while our model fails to surpass the baseline models in terms of accuracy for the balanced case (and those close to it),
we match or exceed the baseline.
%as $\eta$ moves the dataset towards a more imbalanced setting.
In terms of fairness metrics, our approach generally outperforms the baseline models for both of $\eta$.
Detailed results can be found in the Appendix~\ref{ssec:detailed-adult}.

@@

% \begin{figure}
%   \centering
%   \includegraphics[width=0.7\textwidth]{./Figures/nosinn_adult_multiplot_mini_diff.pdf}
%   \caption{
    %   Results for the \textbf{Adult} dataset.
    %   The $x$-axis corresponds to the difference in positive rates.
    %   An ideal result would occupy the \textbf{top-left}.
%   }%
%   \label{fig:adult-chart}
% \end{figure}

>>>>

\paragraph{cMNIST.}
The coloured MNIST (cMNIST) dataset is a variant of the MNIST dataset in which the digits are coloured.
In the training set, the colours have a one-to-one correspondence with the digit class.
In the test set (and the representative set), colours are assigned randomly.
The colours are drawn from Gaussians with 10 different means.
We follow the colourisation procedure outlined by~\citet{kim2019learning}, with the mean colour values selected so as to be maximally dispersed.
The full list of such values can be found in Appendix~\ref{sec:color-details}.
We produce multiple variants of the cMNIST dataset corresponding to different standard deviations $\sigma$ for the colour sampling:
$\sigma \in \{0.00, 0.01, ..., 0.05 \}$.
% The images are symmetrically zero-padded to a size of $32 \times 32$ to allow for an additional downsampling stage in the cFlow model.
% No additional data augmentation is used.
% 24,000 of the 36,000 training samples were set aside for representative, the remaining 10,000 for training the downstream classifier.
% We continue to use 10,000 samples usually reserved for testing on vanilla MNIST for that same purpose.

@@

\paragraph{CelebA.}
To evaluate the effectiveness of our framework on real-world image data we use the CelebA dataset~\citep{liu2015faceattributes}, consisting of 202,599 celebrity images.
These images are annotated with various binary physical  attributes, including ``gender'', ``hair color'', ``young'', etc, from which we  select our sensitive and target attributes.
The images are centre cropped and resized to $64\times64$, as is standard practice.
For our experiments, we designate ``gender'' as the sensitive attribute,
and ``smiling'' and ``high cheekbones'' as target attributes.
We chose gender as the sensitive attribute as it a common sensitive attribute in the fairness literature.
For the target attributes, we chose attributes that are harder to learn than gender and which do not correlate too strongly with gender in the dataset
(``wearing lipstick'' for example being an attribute too closely correlated with gender).
% The sampling and evaluation procedure is identical to that conducted for the UCI Adult dataset.
% We evaluate the performance of all models for mixing factors ($\eta$) of value $\{0, 0.1, ..., 1\}$. 
The model is trained on the representative set (normal subset of CelebA)
and is then used to encode the artificially biased training set and the test set.
The results for the most strongly biased training set ($\eta=0$) can be found in Fig.~\ref{fig:celeba-targets}.
Our method outperforms the baselines in accuracy and fairness.

We also assess performance for different mixing factors ($\eta$) which correspond to varying degrees of bias in the training set
(see Fig.~\ref{fig:celeba-multiplot}).
This is to verify that the model does not \emph{harm} performance when there is not much bias in the training set.
For these experiments, the model is trained once on the representative set and is then used to encode different training sets.
The results show that for the intermediate values of $\eta$, our model incurs a small penalty in terms of accuracy,
but at the same time makes the results \emph{fairer} (corresponding to an accuracy-fairness trade-off). Qualitative results can be found in Fig.~\ref{fig:celeba_cflow} (images from cVAE can be found in section~\ref{sec:qual-results-celeba}).

To show that our method can handle multinomial, as well as binary, sensitive attributes, we also conduct experiments with $s=\textrm{hair color}$ as a ternary attribute (``Blonde'', ``Black'', ``Brown''), excluding ``Red'' because of the paucity of samples and the noisiness of their labels. The results for these experiments can be found in section~\ref{sec:additional-results}.
% Looking at the produced invariant representations,
% The failure of the cVAE to model complex interactions between $x_u$ and $s$ is brought to bear here.
% Due to the balancing between the adversarial and reconstruction losses,
% information relevant to the downstream task of predicting ``smiling'' is lost during the pre-training phase,
% leading to poor performance at all values of $\eta$ except the most extreme ones,
% for which it does reasonably well at mitigating the sampling bias (Fig.~\ref{fig:celeba_chart}).
% Meanwhile, cFlow does not suffer this drawback -- the images remain sharp
% (see Fig.~\ref{fig:cflow_celeba_recon_y} in comparison to Fig. \ref{fig:cvae_celeba_recon_y})
% and the accuracy obtained by training on them is comparable to the baseline at intermediary values -- and at the same time matches and outperforms the cVAE at $\eta = 0$ and $\eta = 1$, respectively.
% For these same values,
% incorporating the mutual information loss proposed by~\cite{kim2019learning} into the classifier's loss function greatly harmed performance measured by accuracy;
% the DP is zero due to all predictions being negative.

@@

\paragraph{Results for the UCI Adult dataset.}
The UCI Adult dataset consists of census data and is commonly used to evaluate models focused on algorithmic fairness.
Following convention, we designate ``gender'' as the sensitive attribute $s$ and whether an individual's salary is \$50,000 or greater as $y$.
We show the performance of our approach in comparison to baseline approaches in Fig. \ref{fig:adult-chart}.
We evaluate the performance of all models for mixing factors ($\eta$) $0$ and $1$. 
Results shown in Fig. \ref{fig:adult-chart} show that
%while our model fails to surpass the baseline models in terms of accuracy for the balanced case (and those close to it),
we match or exceed the baseline.
%as $\eta$ moves the dataset towards a more imbalanced setting.
In terms of fairness metrics, our approach generally outperforms the baseline models for both of $\eta$.
Detailed results can be found in the Appendix~\ref{sec:additional-results}.

@@

% \begin{figure}
%   \centering
%   \includegraphics[width=0.7\textwidth]{./Figures/nosinn_adult_multiplot_mini_diff.pdf}
%   \caption{
    %   Results for the \textsc{Adult} dataset.
    %   The $x$-axis corresponds to the difference in positive rates.
    %   An ideal result would occupy the \textsc{top-left}.
%   }%
%   \label{fig:adult-chart}
% \end{figure}
